---
title: "Homework 1: RMarkdown"
author: "Grace MacKay"
date: Created on "1/24/2020" Submitted on `r format(Sys.time(), "%d %B, %Y")`
output: 
  html_document:
    theme: journal
    toc: TRUE
    toc_float: TRUE
  
---

```{r loading files, echo=FALSE}
## install.packages('gapminder')
## install.packages('DT')
## install.packages('tibble')

library("gapminder")
library("DT")
library("tibble")
```


# Calling the "iris" dataset to view
```{r viewing dataset, echo=FALSE}
iris
```

# Exploring the basics of the "iris" dataset

This dataset is pre-esxisting in R. A list of all of the preloaded datasets in R can be found by running the command data() in the R console.

## Layout of the data

### How is it formatted?
The "iris" data is formatted as a _dataframe_ with **150 observations** and **5 variables** of interest
```{r looking at the structure, echo=FALSE}
str(iris)
```

### What are the names of the variables measured?

The variables of interest are **Sepal Length**, **Sepal Width**, **Petal Length**, **Petal Width** and **Species**. The first four columns have _numeric_ data, providing measurements in cm. The species column contains data characterized as _strings_, which give the names of the species of irises. 
``` {r Column names , echo=FALSE}
names(iris)
```

There are only three species included in the dataset: _setosa_, _versicolor_ and _virginica_.
``` {r flower names, echo=FALSE}
unique(iris$Species)
```


# Summary of the data

```{r summary of whole dataframe, echo=FALSE}
summary(iris)
```


## What is the average iris like?
Average sepal length is **5.843cm**.
```{r sepal length mean, echo=FALSE}
mean(iris$Sepal.Length)
```

Average sepal width is **3.057cm**.
```{r sepal width mean, echo=FALSE}
mean(iris$Sepal.Width)
```

Average petal length is **3.758cm**.
```{r petal length mean, echo=FALSE}
mean(iris$Petal.Length)
```

Average petal width is **1.199cm**.
```{r petal width mean, echo=FALSE}
mean(iris$Petal.Width)
```

## What are the ranges for each variable?

The range for sepal length is **4.3cm to 7.9cm**.
```{r range for sepal length , echo=FALSE}
range(iris$Sepal.Length)
```

The range for sepal width is **2.0cm to 4.4cm**.
```{r range for sepal width , echo=FALSE}
range(iris$Sepal.Width)
```

The range for petal length is **1.0cm to 6.9cm**.
```{r range for petal length , echo=FALSE}
range(iris$Petal.Length)
```

The range for petal width is **1.0cm to 2.5cm**.
```{r range for petal width , echo=FALSE}
range(iris$Petal.Width)
```



## How many of each type of iris was studied?
There were **50** setosa irises, **50** versicolor irises and **50** virginica irises examined.
```{r how many of each iris, echo=FALSE}
summary(iris$Species)
```

## Summaries for each species

### What is the setosa species like?
Summary of the _setosa_ species:
``` {r setosa summary, echo=FALSE}
setosa_only <- subset.data.frame(iris, subset = (iris$Species == "setosa"))
summary(setosa_only)
```
The average setosa sepal length is **5.006cm**.
``` {r setosa sepal length, echo=FALSE}
mean(setosa_only$Sepal.Length)
```

The average setosa sepal width is **3.428cm**.
``` {r setosa sepal width, echo=FALSE}
mean(setosa_only$Sepal.Width)
```

The average setosa petal length is **1.462cm**.
``` {r setosa petal length, echo=FALSE}
mean(setosa_only$Petal.Length)
```

The average setosa petal width is **0.246cm**.
``` {r setosa petal width, echo=FALSE}
mean(setosa_only$Petal.Width)
```

### What is the versicolor species like?

Summary of the _versicolor_ species:
``` {r versicolor summary, echo=FALSE}
versicolor_only <- subset.data.frame(iris, subset = (iris$Species == "versicolor"))
summary(versicolor_only)
```

The average versicolor sepal length is **5.936cm**.
``` {r versicolor sepal length, echo=FALSE}
mean(versicolor_only$Sepal.Length)
```

The average versicolor sepal width is **2.770cm**.
``` {r versicolor sepal width, echo=FALSE}
mean(versicolor_only$Sepal.Width)
```

The average versicolor petal length is **4.620cm**.
``` {r versicolor petal length, echo=FALSE}
mean(versicolor_only$Petal.Length)
```

The average versicolor petal width is **1.326cm**.
``` {r versicolor petal width, echo=FALSE}
mean(versicolor_only$Petal.Width)
```

### What is the virginica species like?

Summary of the _virginica_ species:
``` {r virginica summary, echo=FALSE}
virginica_only <- subset.data.frame(iris, subset = (iris$Species == "virginica"))
summary(virginica_only)
```

The average virginica sepal length is **6.588cm**.
``` {r virginica sepal length, echo=FALSE}
mean(virginica_only$Sepal.Length)
```

The average virginica sepal width is **2.974cm**.
``` {r virginica sepal width, echo=FALSE}
mean(virginica_only$Sepal.Width)
```

The average virginica petal length is **5.552cm**.
``` {r virginica petal length, echo=FALSE}
mean(virginica_only$Petal.Length)
```

The average virginica petal width is **2.026cm**.
``` {r virginica petal width, echo=FALSE}
mean(virginica_only$Petal.Width)
```










```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



